---
title: "SDS315 HW3"
author: "Preston Bui"
date: "2024-02-04"
output:
  html_document:
    df_print: paged
---

EID: ppb434

Github Repo: [](https://github.com/PrestonBui/SDS315-HW3)

```{r message = FALSE, echo = FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(broom)
library(kableExtra)
```

# **Problem 1**
```{r echo = FALSE}
creatinine = read.csv("~/Downloads/creatinine.csv")

ggplot(creatinine, aes(x = age, y = creatclear)) + geom_point() + geom_smooth(method = "lm", se = FALSE) + labs(title = "Creatinine Clearance vs Age", x = "Age (years)", y = "Creatinine Clearance (mL/minute)")

model <- lm(creatclear ~ age, data = creatinine)

prediction <- round(predict(model, newdata = data.frame(age = 55)), 2)
```

## Part A


We should predict a creatinine clearance rate of approximately `r prediction` mL/minute for 55 year olds. This number was determined by using a linear regression model, with the formula Creatinine Clearance = (Slope × Age) + Intercept.

## Part B
```{r echo = FALSE}
rateOfChange <- round(coef(model)["age"], 2)
```

The rate of change of the creatinine clearance rate with age was determined to be approximately `r rateOfChange` mL/minute per year, as given by the slope of the linear regression line. This was calculated using the linear regression model Creatinine Clearance = (Slope × Age) + Intercept, where the slope represents the average decrease in creatinine clearance per year of age.

## Part C
```{r echo = FALSE}
predicted_40 <- round(predict(model, newdata = data.frame(age = 40)), 2)
predicted_60 <- round(predict(model, newdata = data.frame(age = 60)), 2)

given_rate_40 <- 135
given_rate_60 <- 112
```

The expected creatinine clearance rate for a 40-year-old and a 60-year-old were determined using the regression equation Creatinine Clearance = (Slope × Age) + Intercept. The predicted rate for a 40-year-old is approximately `r predicted_40` mL/minute and for a 60-year-old is approximately `r predicted_60` mL/minute. Comparing these to the actual given rates of 135 mL/minute for the 40-year-old and 112 mL/minute for the 60-year-old, it is found that both individuals have clearance rates higher than expected for their ages, with the values for the 40-year-old being `r given_rate_40 - predicted_40` mL/minute higher and for the 60-year-old being `r given_rate_60 - predicted_60` mL/minute higher than the model predictions.


# **Problem 2**
## Introduction
Beta is a measure of a stock's volatility relative to the overall market. It's calculated using regression analysis, where a stock's returns are compared against the market's returns, typically represented by a benchmark index like the S&P 500. A beta value of 1 implies that the stock's price moves with the market, a value greater than 1 indicates higher volatility, and less than 1 signifies lower volatility. By analyzing historical price movements, beta provides investors with a gauge for potential risk and return, serving as a key component in the Capital Asset Pricing Model (CAPM) to assess the expected return on an investment given its risk.

## Table

```{r echo = FALSE}
market <- read.csv("~/Downloads/marketmodel.csv")

results_tibble <- tibble(
  Ticker = character(),
  Intercept = numeric(),
  Beta = numeric(),
  R_squared = numeric(),
  stringsAsFactors = FALSE
)

# List of stock tickers
stocks <- c("AAPL", "GOOG", "MRK", "JNJ", "WMT", "TGT")

# Perform regression for each stock and store the results
for (stock in stocks) {
  # Run the linear regression
  model <- lm(reformulate('SPY', response = stock), data = market)
  summary_model <- summary(model)
  
  # Store the results
  results_tibble <- rbind(results_tibble, tibble(
    Ticker = stock,
    Intercept = coef(summary_model)[1, 1],
    Beta = coef(summary_model)[2, 1],
    R_squared = summary_model$r.squared
  ))
}

# Display the results with kable
results_table <- kable(results_tibble, "html", align = c('l', 'r', 'r', 'r'), 
                       caption = "Estimated Betas for Each Stock",
                       digits = 2) %>% # Set the number of digits here
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "left") %>%
  column_spec(1, bold = TRUE)
# To render the table in the R Markdown document
results_table


```

## Conclusion
In the provided analysis, WMT (Walmart) exhibits the lowest systematic risk with a beta of 0.52, suggesting its stock price is less volatile and thus less sensitive to market swings. On the other end, AAPL (Apple) carries the highest systematic risk with a beta of 1.07, indicating its stock price movements are more closely aligned with, and potentially more reactive to, the overall market trends. These beta values are instrumental in gauging the inherent market risk each stock carries, with lower betas signifying more stability and higher betas signaling greater sensitivity to market changes.


# **Problem 3**
```{r echo = FALSE}
covid_data <- read.csv("~/Downloads/covid.csv")

italy_data <- covid_data %>% filter(country == "Italy", deaths > 0)
spain_data <- covid_data %>% filter(country == "Spain", deaths > 0)

model_italy <- lm(log(deaths) ~ days_since_first_death, data = italy_data)
model_spain <- lm(log(deaths) ~ days_since_first_death, data = spain_data)

growth_rate_italy <- round(exp(coef(model_italy)["days_since_first_death"]), 3)
growth_rate_spain <- round(exp(coef(model_spain)["days_since_first_death"]), 3)

doubling_time_italy <- round(log(2) / coef(model_italy)["days_since_first_death"])
doubling_time_spain <- round(log(2) / coef(model_spain)["days_since_first_death"])

italy_data$Predicted <- exp(predict(model_italy))
spain_data$Predicted <- exp(predict(model_spain))
plot_data <- rbind(
  italy_data %>% mutate(Country = "Italy"),
  spain_data %>% mutate(Country = "Spain")
)

ggplot(plot_data, aes(x = days_since_first_death, y = deaths, color = Country)) + geom_line(aes(y = Predicted), linetype = "dashed") + geom_point(aes(color = Country)) + scale_y_log10() + labs(title = "COVID-19 Deaths Over Time", x = "Days Since First Death", y = "Reported Deaths (log scale)", color = "Country")

```

For Italy, the estimated daily growth rate in COVID-19 deaths is `r growth_rate_italy`, with a doubling time of approximately `r doubling_time_italy` days. Similarly, Spain's estimated growth rate is `r growth_rate_spain`, with a doubling time of `r doubling_time_spain` days. These growth rates reflect how quickly the number of deaths was increasing per day during the initial phase of the pandemic in each country.

The graph above displays the trajectory of reported daily deaths over time for both Italy and Spain, using the number of days since the first death as the time variable. The actual reported deaths are shown with points, and the estimated exponential growth curves are shown with dashed lines, each distinguished by color for clarity between the countries.



# **Problem 4**
```{r echo = FALSE}
milk <- read.csv("~/Downloads/milk.csv")

milk <- milk %>%
  mutate(log_price = log(price), log_sales = log(sales))

model <- lm(log_sales ~ log_price, data = milk)

price_elasticity <- coef((model)["log_price"])
```

Using a log-log regression model, we estimated the price elasticity of demand for milk. By regressing the natural log of sales on the natural log of prices from the dataset, the slope of the regression line `r round(coef(model)["log_price"], 3)` provides the elasticity estimate. This value indicates that for a 1% increase in the price of milk, the quantity demanded changes by `r round(coef(model)["log_price"], 3)`%. The results suggest that milk demand is elastic based on the absolute value of elasticity being greater than 1.
